Static Discharge of Dynamic, Higher-Order Contracts

### 2013-11-15 pldi2014/content/calculus.tex
Formatting semantics.

\newcommand{\ks}[0]{\gamma^*}
\newcommand{\vv}[0]{\mathbf{v}}
\newcommand{\vvp}[0]{\mathbf{v'}}
\newcommand{\sexp}[4]{\mathbf{eval}(#1,#2,#3,#4)}
\newcommand{\sval}[3]{\mathbf{value}(#1,#2,#3)}
\newcommand{\sapp}[4]{\mathbf{apply}(#1,#2,\ell,#3,#4)}
\newcommand{\simpz}[6]{\mathbf{imp_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\simpo}[5]{\mathbf{imp_1}(#1,#2,#3,#4,#5)}
\newcommand{\schaz}[6]{\mathbf{chap_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\schao}[5]{\mathbf{chap_1}(#1,#2,#3,#4,#5)}
\newcommand{\serr}[3]{\mathbf{error}(#1,#2,#3)}
\newcommand{\sbla}[2]{\mathrm{blame}_{#1}(#2)}
\newcommand{\app}[4]{\mathrm{apply}(#1,#2,#3,#4)}
\newcommand{\red}[2]{\begin{equation}\begin{aligned}& #1\\\rr\, & #2\end{aligned}\end{equation}}
\newcommand{\reda}[2]{$#1\rr#2$}

@@

@@

\begin{tabular}{ r r l l }
\stxclass{\varsigma}{State}{$\sexp{\ks}{\sigma}{\rho}{e}$}{evaluation}\\
\stxclasscont{$\sapp{\ks}{\sigma}{f}{\vv}$}{apply}\\
\stxclasscont{$\sval{\ks}{\sigma}{\vv}$}{value}\\
\stxclasscont{$\serr{\ks}{\sigma}{b}$}{error}\\
\stxclasscont{$\simpz{\ks}{\sigma}{\ell}{f}{w}{\vv}$}{imp0}\\
\stxclasscont{$\simpo{\ks}{\sigma}{\ell}{w}{\vv}$}{imp1}\\
\stxclasscont{$\schaz{\ks}{\sigma}{\ell}{f}{w}{\vv}$}{chap0}\\
\stxclasscont{$\schao{\ks}{\sigma}{\ell}{w}{\vv}$}{chap1}\\
\stxclass{\gamma}{Frame}{\letk{\vx}{\rho}{e}}{let}\\
\stxclasscont{\impcwk{f}{\vv}}{impersonator negative}\\
\stxclasscont{\impcfk{w}}{impersonator function}\\
\stxclasscont{\impcrk{\vv}}{impersonator positive}\\
\stxclasscont{\chacwk{f}{\vv}}{chaperone negative}\\
\stxclasscont{\chacfk{w}}{chaperone function}\\
\stxclasscont{\chacrk{\vv}}{chaperone positive}\\
\stxclass{v}{Val}{$f\,|\,n\,|\,t$}{values}\\
\stxclass{\sigma}{Store}{\funarr{Addr}{Val}}{stores}\\
\stxclass{\rho}{Env}{\funarr{Var}{Addr}}{environments}\\
\stxclass{f,w}{Op}{\nttwo{clos}{\mathit{lam}}{\rho}}{closures}\\
\stxclasscont{\ntthr{imp}{\ell}{f}{w}}{impersonators}\\
\stxclasscont{\ntthr{chap}{\ell}{f}{w}}{chaperones}\\
\stxclasscont{$p$}{primitives}\\
\stxclass{a}{Addr}{an infinite set of addresses}{}
\end{tabular}

@@

\noindent
\reda{\sexp{\ks}{\sigma}{\rho}{\appe{\ae_f}{\ae_1\,\dots\,\ae_n}}}{\sapp{\ks}{\sigma}{f}{\langle v_1,\dots,v_n\rangle}}
where $f=\A{\sigma}{\rho}{\ae_f}$ and $v_i=\A{\sigma}{\rho}{\ae_i}$ for $i=1,\dots,n$.

@@

\noindent
\reda{\sexp{\ks}{\sigma}{\rho}{\lete{x_1\,\dots\, x_n}{e_0}{e_1}}}{\sexp{\letk{\vx}{\rho}{e_1}::\ks}{\sigma}{\rho}{e_0}} where $\vx=\langle x_1,\dots,x_n\rangle$.

\reda{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma'}{\rho'}{e_1}}
where $(\sigma',\rho')=\bind{\vx}{\vv}$ if $|\vx|=|\vv|$

\reda{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\dynbla{\ks}}
if $|\vx|\ne|\vv|$

@@

\reda{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}{\sexp{\ks}{\sigma}{\rho}{e_c}}
if $v_t\ne\false$
where $v_t=\A{\sigma}{\rho}{\ae_t}$.

\reda{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}{\sexp{\ks}{\sigma}{\rho}{e_a}}
if $v_t=\false$
where $v_t=\A{\sigma}{\rho}{\ae_t}$.
\end{figure*}

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\appe{\ae_f}{\ae_1\,\dots\,\ae_n}}}{\sapp{\ks}{\sigma}{f}{\langle v_1,\dots,v_n\rangle}}

\noindent
where $f=\A{\sigma}{\rho}{\ae_f}$ and $v_i=\A{\sigma}{\rho}{\ae_i}$ for $i=1,\dots,n$.

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\lete{x_1\,\dots\, x_n}{e_0}{e_1}}}{\sexp{\letk{\vx}{\rho}{e_1}::\ks}{\sigma}{\rho}{e_0}}

\noindent
where $\vx=\langle x_1,\dots,x_n\rangle$.

\red{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma'}{\rho'}{e_1}}

\noindent
where $(\sigma',\rho')=\bind{\vx}{\vv}$ if $|\vx|=|\vv|$

\red{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\dynbla{\ks}} if $|\vx|\ne|\vv|$

@@

\red{\sapp{\ks}{\sigma}{\clo{\lame{x_1\,\dots\,x_n}{e}}{\rho}}{\vv}}{\sexp{\ks}{\sigma'}{\rho'}{e}}
if $|\vx|=|\vv|$ where $\vx=\langle x_1,\dots,x_n\rangle$ and $(\sigma',\rho')=\bind{\vx}{\vv}$

@@

\red{\sapp{\ks}{\sigma}{\clo{\lame{x_1\,\dots\,x_n}{e}}{\rho}}{\vv}}{\dynbla{\ks}}
if $|\vx|\ne|\vv|$ where $\vx=\langle x_1,\dots,x_n\rangle$

@@

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sapp{\ks}{\sigma}{f}{\vvp}}

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\schaz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}
if $|\vv|=|\vvp|$ and $\mathrm{operator?}(w)$.

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\sbla{3}{\ell}}
if $|\vv|=|\vvp|$ but $\mathrm{operator?}(w)$ does not hold.

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sbla{4}{\ell}}
if $|\vv|\ne|\vvp|$ and $|\vv|+1\ne|\vvp|$.

\red{\sval{\impcrk{\vv}::\ks}{\sigma}{\vvp}}{\sval{\ks}{\sigma}{\vvp}}
if $|\vv|=|\vvp|$

\red{\sval{\impcrk{\vv}::\ks}{\sigma}{\vvp}}{\serr{\ks}{\sigma}{?}}
if $|\vv|\ne|\vvp|$

@@

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sapp{\ks}{\sigma}{f}{\vvp}}
if $|\vv|=|\vvp|$ and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sbla{1}{\ell}}
if $|\vv|=|\vvp|$ but $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$ does not hold.

--

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\schaz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}
if $|\vv|=|\vvp|$, $\mathrm{operator?}(w)$, and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\sbla{3}{\ell}}
if $|\vv|=|\vvp|$ but $\mathrm{operator?}(w)$ does not hold.

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\sbla{1}{\ell}}
if $|\vv|=|\vvp|$ and $\mathrm{operator?}(w)$ but $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$ does not hold.

--

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sbla{4}{\ell}}
if $|\vv|\ne|\vvp|$ and $|\vv|+1\ne|\vvp|$

\red{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}{\sval{\ks}{\sigma}{\vvp}}
if $|\vv|=|\vvp|$ and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

\red{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}{\sbla{7}{\ell}}
if $|\vv|=|\vvp|$ but $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$ does not hold

\red{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}{\sbla{6}{\ell}}
if $|\vv|\ne|\vvp|$

\red{\sapp{\ks}{\sigma}{p}{\vv}}{\serr{\ks}{\sigma}{\delta(p,\vv)}}
if $\delta(p,\vv)=b$ for some $b$.

>>>>

\newcommand{\ks}[0]{\gamma^*}
\newcommand{\vv}[0]{\mathbf{v}}
\newcommand{\vvp}[0]{\mathbf{v'}}
\newcommand{\sexp}[4]{\mathbf{eval}(#1,#2,#3,#4)}
\newcommand{\sval}[3]{\mathbf{value}(#1,#2,#3)}
\newcommand{\sapp}[4]{\mathbf{apply}(#1,#2,\ell,#3,#4)}
\newcommand{\simpz}[6]{\mathbf{imp_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\simpo}[5]{\mathbf{imp_1}(#1,#2,#3,#4,#5)}
\newcommand{\schaz}[6]{\mathbf{chap_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\schao}[5]{\mathbf{chap_1}(#1,#2,#3,#4,#5)}
\newcommand{\serr}[3]{\mathbf{error}(#1,#2,#3)}
\newcommand{\sbla}[2]{\mathrm{blame}_{#1}(#2)}
\newcommand{\app}[4]{\mathrm{apply}(#1,#2,#3,#4)}
\newcommand{\reda}[2]{\begin{equation}\begin{aligned}& #1\\\rr\, & #2\end{aligned}\end{equation}}
\newcommand{\red}[2]{$#1\rr#2$}
\newcommand{\redtwo}[5]{\begin{math}#1\rr\begin{cases}#2 & #3\\ #4 & #5\end{cases}\end{math}}
\newcommand{\redthr}[7]{\begin{math}#1\rr\begin{cases}#2 & #3\\ #4 & #5\\#6 & #7\end{cases}\end{math}}

@@

\newcommand{\blame}[1]{\ensuremath{\mathrm{blame}(#1)}}
\newcommand{\impblame}[1]{\ensuremath{\mathrm{imp\mhyphen blame}(#1)}}
\newcommand{\chablame}[1]{\ensuremath{\mathrm{cha\mhyphen blame}(#1)}}
\newcommand{\noblame}[0]{\ensuremath{\mathrm{noblame}}}

@@

\begin{tabular}{ r r l l }
\stxclass{\varsigma}{State}{$\sexp{\ks}{\sigma}{\rho}{e}$}{evaluation}\\
\stxclasscont{$\sapp{\ks}{\sigma}{f}{\vv}$}{apply}\\
\stxclasscont{$\sval{\ks}{\sigma}{\vv}$}{value}\\
\stxclasscont{$\serr{\ks}{\sigma}{b}$}{error}\\
\stxclasscont{$\simpz{\ks}{\sigma}{\ell}{f}{w}{\vv}$}{imp0}\\
\stxclasscont{$\simpo{\ks}{\sigma}{\ell}{w}{\vv}$}{imp1}\\
\stxclasscont{$\schaz{\ks}{\sigma}{\ell}{f}{w}{\vv}$}{chap0}\\
\stxclasscont{$\schao{\ks}{\sigma}{\ell}{w}{\vv}$}{chap1}\\
\stxclass{\gamma}{Frame}{\letk{\vx}{\rho}{e}}{let}\\
\stxclasscont{\impcwk{f}{\vv}}{negative guard}\\
\stxclasscont{\impcfk{w}}{impersonated procedure}\\
\stxclasscont{\impcrk{\vv}}{positive guard}\\
\stxclasscont{\chacwk{f}{\vv}}{chaperone negative}\\
\stxclasscont{\chacfk{w}}{chaperone function}\\
\stxclasscont{\chacrk{\vv}}{chaperone positive}\\
\stxclass{v}{Val}{$f\,|\,n\,|\,t$}{values}\\
\stxclass{\sigma}{Store}{\funarr{Addr}{Val}}{stores}\\
\stxclass{\rho}{Env}{\funarr{Var}{Addr}}{environments}\\
\stxclass{f,w}{Op}{\nttwo{clos}{\mathit{lam}}{\rho}}{closures}\\
\stxclasscont{\ntthr{imp}{\ell}{f}{w}}{impersonators}\\
\stxclasscont{\ntthr{chap}{\ell}{f}{w}}{chaperones}\\
\stxclasscont{$p$}{primitives}\\
\stxclass{a}{Addr}{an infinite set of addresses}{}\\
\stxclass{b}{Blame}{$\blame{\ell}\,|\,\noblame$}{blame}
\end{tabular}

@@

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\appe{\ae_f}{\ae_1\,\dots\,\ae_n}}}{\sapp{\ks}{\sigma}{f}{\langle v_1,\dots,v_n\rangle}}
where $f=\A{\sigma}{\rho}{\ae_f}$ and $v_i=\A{\sigma}{\rho}{\ae_i}$ for $i=1,\dots,n$.

@@

% encounter let form
\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\lete{x_1\,\dots\, x_n}{c}{e}}}{\sexp{\letk{\vx}{\rho}{e}::\ks}{\sigma}{\rho}{c}}
where $\vx=\langle x_1,\dots,x_n\rangle$.

\redtwo{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}
       {\sexp{\ks}{\sigma'}{\rho'}{e_1}}{\mbox{if }|\vx|=|\vv|}
       {\serr{\ks}{\sigma}{\noblame}}{\mbox{if }|\vx|\ne|\vv|}
where $(\sigma',\rho')=\bind{\vx}{\vv}$

\redtwo{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}
       {\sexp{\ks}{\sigma}{\rho}{e_c}}{\mbox{if }v_t\ne\false}
       {\sexp{\ks}{\sigma}{\rho}{e_a}}{\mbox{if }v_t=\false}
where $v_t=\A{\sigma}{\rho}{\ae_t}$.

\redtwo{\sapp{\ks}{\sigma}{\clo{\lame{x_1\,\dots\,x_n}{e}}{\rho}}{\vv}}
       {\sexp{\ks}{\sigma'}{\rho'}{e}}{\mbox{if }|\vx|=|\vv|}
       {\serr{\ks}{\sigma}{\blame{\ell}}}{\mbox{if }|\vx|\ne|\vv|}
where $\vx=\langle x_1,\dots,x_n\rangle$ and $(\sigma',\rho')=\bind{\vx}{\vv}$

\redtwo{\sval{\impcwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}
       {\simpz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\mbox{if }|\vv|=|\vvp|\mbox{ and } \mathrm{operator?}(w)\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\mathrm{operator?}(w)\mbox{.}}

\redtwo{\sval{\impcwk{f}{\vv}::\ks}{\sigma}{\vvp}}
       {\sapp{\ks}{\sigma}{f}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{.}}

\redtwo{\sval{\impcrk{\vv}::\ks}{\sigma}{\vvp}}
       {\sval{\ks}{\sigma}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{ and }|\vv|+1\ne|\vvp|\mbox{.}}

\redthr{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}
       {\schaz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\mbox{if }|\vv|=|\vvp|\mbox{, } \mathrm{operator?}(w)\mbox{, and }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\mathrm{operator?}(w)\mbox{,}}
       {\serr{\ks}{\sigma}{\chablame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ and }\mathrm{operator?}(w)\mbox{ but not }\chapof{\vvp}{\vv}\mbox{.}}

\redthr{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}
       {\sapp{\ks}{\sigma}{f}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{ and }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\chablame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{ and }|\vv|+1\ne|\vvp|\mbox{.}}

\redthr{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}
       {\sval{\ks}{\sigma}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{ and }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\chablame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{.}}

\red{\sapp{\ks}{\sigma}{p}{\vv}}{\serr{\ks}{\sigma}{\delta(p,\vv)}}
if $\delta(p,\vv)=b$ for some $b$.
\end{figure*}

@@

@@

@@

@@

@@
